The \dealii Library, Version 9.2

### 2020-05-14 9.2/paper.tex

The major changes of this release are:
%
\begin{itemize}
\item
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}),
%   \item Dedicated support for symbolic algebra (see
%     Section~\ref{subsec:sd}),
%   \item Full support for $hp$~adaptivity in parallel computations (see
%     Section~\ref{subsec:hp}),
%   \item An interface to the HDF5 file format and libraries (see
%     Section~\ref{subsec:hdf5}),
%   \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
%   \item Parallel geometric multigrid (GMG) improvements (see
%     \cite{ClevengerHeisterKanschatKronbichler2019} and
%     Section~\ref{subsec:gmg}),
%   \item Four new tutorial programs (step-61, step-62, step-63, step-64),
%     as well as one new code gallery program (see
%     Section~\ref{subsec:steps}).
\end{itemize}
%
The major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item
% \item
%   The release contains a number of performance improvements and bug fixes for
%   the matrix-free framework. One notable improvement is the support for
%   renumbering of degrees of freedom within the cells for discontinuous
%   elements, avoiding some reshuffling operations across the SIMD lanes
%   with vectorization over several cells and faces, which is especially
%   useful on processors with AVX-512 vectorization (8 doubles), speeding up
%   operations by up to 10\%. Secondly, the strategy for the most efficient
%   tensor product evaluators according to the performance analysis of
%   \cite{KronbichlerKormann2019} in the context of more quadrature points than
%   shape functions has been revised for better performance.
%
% \item A new class \texttt{ParsedConvergenceTable} has been introduced
%   that greatly simplifies the construction of convergence tables,
%   reading the options for the generation of the table from a parameter
%   file, and providing methods that, combined with a parameter file,
%   allow one to generate convergence tables using one-liners in user
%   codes.
%
% \item
%   The \texttt{FE\_BernardiRaugel} class implements the non-standard
%   Bernardi-Raugel (BR) element that can be used to construct a stable
%   velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
%   element is an enriched version of the $Q_1^d$ element with added bubble
%   functions on each edge (in 2d) or face (in 3d). It addresses the fact
%   that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
%   larger velocity space), and that the $Q_2^d\times Q_0$ combination is
%   stable but converges with only first-order at the cost of the large
%   number of velocity unknowns. The BR space is thus intermediate between the
%   $Q_1^d$ and $Q_2^d$ spaces.
%
%   The element is currently only implemented for parallelogram meshes due to
%   difficulties associated with the mapping of shape functions: The shape
%   functions of the $Q_1^d$ part of the element need to be mapped as
%   scalars, as is common for the vector components of the $Q_1^d$ element;
%   on the other hand, the vector-valued edge bubble functions need to be
%   mapped using the Piola transform as is common for the Raviart-Thomas
%   element. \dealii{} does not currently have the ability to use different
%   mappings for individual shape functions, though this functionality is
%   planned for the next release.
%
% \item
%   The \texttt{FE\_NedelecSZ} class is a new implementation of the
%   N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
%   the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
%   present in traditional N{\'e}d{\'e}lec elements that arise from the edge
%   and face parameterizations used in the basis functions. Therefore, this
%   element should provide consistent results for general quadrilateral and
%   hexahedral elements for which the relative orientations of edges and
%   faces (as seen from all adjacent cells) are often difficult to establish.
%   The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
%   assigning a globally defined orientation to local edges and faces. A
%   detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
%   element in \dealii{} can be found in \cite{Kynch2017}.
%
% \item All of the elementary geometrical objects of the library (namely
%   \texttt{Point<dim>}, \texttt{Segment<dim>}, and
%   \texttt{BoundingBox<dim>}) have been augmented with the traits
%   needed to comply with \texttt{boost::geometry} concepts. A new
%   interface to \texttt{boost::geometry::index::rtree} has been added
%   that simplifies the construction of spatial indices based on points,
%   bounding boxes, or segments.
\end{itemize}
%
In addition to these changes, the changelog lists more than 200 other
features and bugfixes.

@@

Many of the \dealii{} tutorial programs were substantially revised as
part of this release. In particular, we have converted many places
that now allow for simpler code through the use of C++11 features such
as range-based for loops and lambda functions.

In addition, there are seven new tutorial programs:
\begin{itemize}
\item \texttt{step-47}
\item \texttt{step-50}
\item \texttt{step-58}
\item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
manifold class that can propagate curved boundary information into the
interior of a computational domain, and \texttt{MappingQCache}, which can sample 
the information of expensive manifolds in the points of a \texttt{MappingQ} and 
cache it for further use.
\item \texttt{step-67} presents an explicit time integrator for the
compressible Euler equations discretized with a high-order discontinuous
Galerkin scheme using the matrix-free infrastructure. Besides the use of
matrix-free evaluators for systems of equations and over-integration, it also
presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
of the action of the inverse mass matrix in the DG setting using tensor
products. Furthermore, this tutorial demonstrates the usage of new 
pre and post operations, which can be passed to \texttt{cell\_loop()}, to schedule operations on sections of vectors close 
to the matrix-vector product to increase data locality 
and discusses performance-related aspects.
\item \texttt{step-69}
\item \texttt{step-70}
\end{itemize}

@@

The 9.2 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
     {around 15 incompatible changes}; see \cite{changes92}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions; and the majority change internal
interfaces that are not usually used in external
applications. However, some are worth mentioning:
\begin{itemize}
\item The functions:
\begin{itemize}
\item \texttt{DoFHandler::loccaly\_owned\_dofs\_per\_processor()}
\item \texttt{DoFHandler::loccaly\_owned\_mg\_dofs\_per\_processor()}
\end{itemize} have been deprecated. As discussed in Subsection~\ref{subsec:performance}, deal.II does not store information for all processes on all processes processes, but only the local information or the locally-relevant information. Users are asked to construct the global information on their own, e.g. by calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
\item

>>>>

The major changes of this release are:
%
\begin{itemize}
\item xy \todo[inline]{Update once we have the subsections in Section 2;
  provide cross-references to each of these subsections}
%   \item Improved support for automatic differentiation (see
%     Section~\ref{subsec:ad}), 
\end{itemize}
%
The major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
\item x \todo[inline]{Wolfgang to write about complex-valued output}
\item y \todo[inline]{Timo to write about problems and fixes for
  computations with more than $2^{32}$ unknowns}
\item z \todo[inline]{What else? Maybe mention the updated step-12?}
\end{itemize}
%
In addition to these changes, the changelog lists more than 200 other
features and bugfixes.

@@

Many of the \dealii{} tutorial programs were revised in a variety of
ways as part of this release. A particular example is that we have
converted a number of programs to use range-based for loops (a C++11
feature) for loops over a range of integer indices such as loops over
all quadrature points or all indices of degrees of freedom during
assembly. This makes sense given that the
range-based way of writing loops seems to be the idiomatic approach
these days, and that we had previously already converted loops over
all cells in this way.

In addition, there are a number of new tutorial programs:
\begin{itemize}
\item \texttt{step-47}
\todo[inline]{Zhuoran to write}
\item \texttt{step-50}
\todo[inline]{Timo/Conrad/... to write}
\item \texttt{step-58}
\todo[inline]{Wolfgang to write}  
\item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
manifold class that can propagate curved boundary information into the
interior of a computational domain, and \texttt{MappingQCache}, which can sample 
the information of expensive manifolds in the points of a \texttt{MappingQ} and 
cache it for further use.
\item \texttt{step-67} presents an explicit time integrator for the
compressible Euler equations discretized with a high-order discontinuous
Galerkin scheme using the matrix-free infrastructure. Besides the use of
matrix-free evaluators for systems of equations and over-integration, it also
presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
of the action of the inverse mass matrix in the DG setting using tensor
products. Furthermore, this tutorial demonstrates the usage of new 
pre and post operations, which can be passed to \texttt{cell\_loop()}, to schedule operations on sections of vectors close 
to the matrix-vector product to increase data locality 
and discusses performance-related aspects.
\item \texttt{step-69}
\todo[inline]{Matthias/Ignacio to write}  
\item \texttt{step-70}
  \todo[inline]{Also need to update announce and announce-short if this
    makes it into the release.}
\end{itemize}

\todo[inline]{Do we have new code gallery programs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Support for large, fully distributed meshes}
\label{subsec:pfT}

\todo[inline]{Peter: Write something about p::f::T}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Better support for parallel $hp$-adaptive algorithms}
\label{subsec:hp}

\todo[inline]{Marc: Your section}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Support for particle-based methods}
\label{subsec:particles}

\todo[inline]{Luca: Your section}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Python interfaces}
\label{subsec:python}

\todo[inline]{What's new here? mention step-49 and step-53 versions written in python.}

@@

The 9.2 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
     {around 60 incompatible changes}; see \cite{changes92}. The majority of these changes
should not be visible to typical user codes; some remove previously
deprecated classes and functions; and the majority change internal
interfaces that are not usually used in external
applications. However, some are worth mentioning:
\begin{itemize}
\item The functions:
\begin{itemize}
\item \texttt{DoFHandler::loccaly\_owned\_dofs\_per\_processor()}
\item \texttt{DoFHandler::loccaly\_owned\_mg\_dofs\_per\_processor()}
\end{itemize} have been deprecated. As discussed in Subsection~\ref{subsec:performance}, deal.II does not store information for all processes on all processes processes, but only the local information or the locally-relevant information. Users are asked to construct the global information on their own, e.g. by calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
\item
